---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ODVC

<!-- badges: start -->
[![R-CMD-check](https://github.com/ryancbushman/ODVC/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ryancbushman/ODVC/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ODVC is to provide tools that allow researchers to do optimal design
of experiments for one-way and two-way nested random effects models.

In the optimal design of experiments framework, the goal is to optimize the 
model information matrix. The inverse of the information matrix is the 
covariance matrix for the maximum likelihood estimators (MLE). By maximizing the 
information of an experiment on some optimality criteria, one simultaneously 
minimizes the variance of the maximum likelihood estimators. The information 
matrix or MLE covariance matrix or can be scored using optimality criteria such 
as the $D$ criteria (the determinant) or the $A$ criteria (the trace). The tools
of the ODVC package pose optimization as a minimization problem. In a 
random-effects model, the information matrix is a function of the 
experiment's sample sizes and the values of the variance components. 

## Installation

You can install the development version of ODVC from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ryancbushman/ODVC")
```

## Example

We will walk through three example problems that demonstrate the use of ODVC
tools. The first is the one-way balanced design. The generate_designs_B function
will create a dataset of designs based on all combinations of the provided
arguments. The user can then subset the dataset with subset_designs_B to see 
designs of a specific size. Finally the user can create a dashboard of plots
to compare the performance of the subset of designs.

```{r example}
library(ODVC)

###### one-way balanced designs ######

ex_1_designs <- generate_designs_B(ngroups = c(2, 3, 4, 6, 8, 12), 
                                   nreps = c(2, 3, 4, 6, 8, 12), 
                                   taus = c(1, 2))

ex_1_designs_subset <- subset_designs_B(data = ex_1_designs, N = 24)

compare_designs_B(designs = ex_1_designs_subset, criteria = "A")

```
Once the user has seen the dashboard and decided on an appropriate design, they
can visualize the design using the plot_design function. 

```{r}
plot_design(n = rep(3, 8) , a = 8, sig_a_sq = 1, error_sq = 1, criteria = "A")
```
If the user would like to explore unbalanced one-way designs, they can use the
generate_designs_U function to generate all unbalanced designs of size "N" with 
"a" number of groups and user provided values for \eqn{\sigma^2_A} and 
\eqn{\sigma^2}. The user can then visualize the performance of the generated
datasets using the compare_designs_U function.

```{r}

##### one-way unbalanced designs #####

ex_2_designs <- generate_designs_U(N = 24, a = 3, sig_a_sq = 2, error_sq = 1)

compare_designs_U(data = ex_2_designs, criteria = "A")

```

Once the user selects the desired design, they can visualize it using the 
plot_design function.

```{r}
plot_design(n = c(9, 8, 7), a = 3, sig_a_sq = 2, error_sq = 1, criteria = "A")
```
Due to the combinatorial nature of optimal design for random effects models, it
is too computationally expensive to generate all two-way designs of size "N".
Rather, designs are generated from one of two classes called 
\eqn{\mathcal{C}_{3,2}} and \eqn{\mathcal{C}_{3,3}}. The first subscript refers
to the number of variance components. The second subscript refers to the
maximum number of subgroups and replicates per subgroup. The user can use the 
generate_two_way_designs function to generate all designs of size "N" within the
preferred class using "s" argument to indicate the second subscript. Once the
dataset has been generated, the user can use the compare_designs_U function
to compare performance.

```{r}

##### two_way designs #####

ex_3_designs <- generate_two_way_designs(N = 12, s = 2, sig_a_sq = 0.1,
                                         sig_b_sq = 2, error_sq = 1)

compare_designs_U(ex_3_designs, criteria = "A")
```

Upon choosing the desired experiment design, the user can reference dataset of
designs and provide arguments for the plot_design_2 function.

```{r}
plot_design_2(a = 4, b_i = c(2, 2, 2, 2), n_ij = c(2, 2, 2, 2, 1, 1, 1, 1),
              n_i_dot = c(4, 4, 2, 2), sig_a_sq = 0.1, sig_b_sq = 2, 
              error_sq = 1, criteria = "A", balanced = FALSE)
```
For more information on how this package can be used to solve optimal design
problems for nested random-effects models, please reference the provided 
vignettes and documentation. 
