---
title: "Tools For One-Way Models"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Tools_For_One_Way_Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ODVC)
library(ggplot2)
```

If a user would like to explore a wide variety of balanced one-way nested random
effects designs with many different parameter settings, they can use 
contour_designs_B function.

```{r, fig.show='hide', warning=FALSE}

contour_plots <- contour_designs_B(ngroups = c(2, 10), nreps = c(2, 10), 
                                   taus = 2)

```

```{r}
contour_plots[[1]][[4]]
contour_plots[[2]][[4]]
```
Once the user narrows down what paremeter settings they would like to explore,
they can use the generate_designs_B to generate balanced designs.

```{r}
candidates_B <- generate_designs_B(ngroups = c(2, 3, 4, 6, 8, 12), 
                                   nreps = c(2, 3, 4, 6, 8, 12),
                                   taus = c(0.5, 1))
head(candidates_B)
```

The subset_designs_B function can be used to select only the designs of a
particular size.

```{r}
candidates_B_N24 <- subset_designs_B(data = candidates_B, N = 24)
head(candidates_B_N24)
```
The user can then use the compare_designs_B function to create a dashboard that
will help them to choose the optimal design.

```{r}
compare_designs_B(designs = candidates_B_N24, criteria = "A")
```
From the dashboard, the user can see that when $\sigma^2_A = 0.5$, the best 
design when N=24 under the $A$ criteria is the one that has 6 groups. The user
can now use the plot_design function to visualize the design.

```{r fig.height=2, fig.width=6}
plot_design(n = rep(4, 6), a = 6, sig_a_sq = 0.5, error_sq = 1, criteria = "A")
```

If the user would like to explore unbalanced designs, they can use the 
generate_designs_U function.

```{r}
options(width = 60)
candidates_U <- generate_designs_U(N = 23, a = 8, sig_a_sq = 0.5, error_sq = 1)

head(candidates_U)
```

To compare the performance of these designs they can use the compare_designs_U
function. This generates a scatter plot of relative efficiency and index within
the design dataframe. The user can also set the top_5 argument equal to TRUE to
label only the top 5 performing designs.

```{r}
compare_designs_U(data = candidates_U, criteria = "A", top_5 = TRUE)
```

From the plot, it is evident that the best design is the one with index 146 from
the dataframe. By referencing this row in the dataframe, the user can use the 
plot_design function to visualize this optimal design.

```{r fig.height=2, fig.width=6}
plot_design(n = c(rep(3, 7), 2), a = 8, sig_a_sq = 0.5, error_sq = 1, 
            criteria = "A")
```

In the optimal design of experiments framework, the goal is to optimize the 
model information matrix. The inverse of the information matrix is the 
covariance matrix for the maximum likelihood estimators. By maximizing the 
information of an experiment on some optimality criteria, one simultaneously 
minimizes the variance of the maximum likelihood estimators. The information 
matrix can be scored using optimality criteria such as the $D$ criteria (the 
determinant) or the $A$ criteria (the trace). The tools of the ODVC package 
define the optimal design as the design the minimizes the optimality criteria. 
In a random-effects model, the information matrix is a function of the 
experiment's sample sizes and the values of the variance components. The user 
can use the general_variance_2VC function to generate the covariance matrix of 
the maximum likelihood estimators (the object being optimized) for a particular 
experiment.

```{r}
optimal_design_inv_info_matrix <- general_variance_2VC(N = 23, n = c(rep(3, 7), 2), 
                                                   a = 8, sig_a_sq = 0.5, 
                                                   error_sq = 1)
optimal_design_inv_info_matrix
```
The user can then use either the A_crit or D_crit function to report a score.

```{r}
A_crit(optimal_design_inv_info_matrix)
D_crit(optimal_design_inv_info_matrix)
```

