---
title: "Tools For Two-Way Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tools_For_Two_Way_Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ODVC)
```

Optimal design of experiments for a random-effects model is a combinatorial 
problem in nature. The reason being is that the object being optimized, the
inverse information matrix, is a function of the sample sizes of the experiment 
and the values of the variance components. When searching the space of possible
designs, one must generate all designs of size $N$ and score the inverse
information matrix of each design on some optimality criteria. When exploring
two-way nested models, the amount of possible designs of size $N$ grows 
exceptionally fast as $N$ increases. Two mitigate this, the ODVC generates all
designs of a chosen $N$ from one of two classes, $\mathcal{C}_{3,2}$ and 
$\mathcal{C}_{3,3}$. The first subscript indicates the number of variance 
components. The second subscript indicates the maximum number of subgroups per
group and the maximum number of replicates per subgroup. 

The user can use the generate_two_way_designs function to generate all designs
of size $N$ from class $s$, with variance components $\sigma^2_A$, $\sigma^2_B$,
and $\sigma^2$.

```{r}
candidates <- generate_two_way_designs(N = 12, s = 2, sig_a_sq = 10, 
                                       sig_b_sq = 2, error_sq = 1)
```

Once the dataframe of designs is generated, the user can call the 
compare_designs_U function to compare design performance.

```{r}
compare_designs_U(data = candidates, criteria = "A")
```
The plot indicates that design 64 from the dataframe is the optimal design. By
referencing the dataframe, the user can call the plot_design_2 function to 
visualize the design.

```{r}
plot_design_2(a = 8,
              b_i = c(2, 2, 2, 1, 1, 1, 1, 1),
              n_ij = c(2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
              n_i_dot = c(3, 2, 2, 1, 1, 1, 1, 1),
              sig_a_sq = 10,
              sig_b_sq = 2,
              error_sq = 1,
              balanced = FALSE,
              criteria = "A")
```

We will now walk through a different design scenario and but generate designs 
from $\mathcal{C}_{3,3}$ rather than $\mathcal{C}_{3,2}$. Note that this takes 
much longer to do because the number of possible designs is much greater.

```{r}
candidates2 <- generate_two_way_designs(N = 12, s = 3, sig_a_sq = 10, 
                                       sig_b_sq = 10, error_sq = 1)
```

Once the dataframe of designs is generated, the user can call the 
compare_designs_U function to compare design performance.

```{r}
design_performance <- compare_designs_U(data = candidates2, criteria = "A")
design_performance + ylim(90, 100)
```

The plot indicates that design 233 from the dataframe is the optimal design. By
referencing the dataframe, the user can provide the necessary information to 
call the plot_design_2 function and visualize the design.

```{r}
plot_design_2(a = 4,
              b_i = c(3, 3, 3, 2),
              n_ij = c(2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
              n_i_dot = c(4, 3, 3, 2),
              sig_a_sq = 10,
              sig_b_sq = 10,
              error_sq = 1,
              balanced = FALSE,
              criteria = "A")
```

The tools in the ODVC package score the designs for the user. If the user would 
like to generate the actual matrix that is being optimized themselves, they can 
call the general_variance_3VC function.

```{r}
optimal_design <- general_variance_3VC(N = 12, n_i_dot = c(4, 3, 3, 2), 
                     n_ij = c(2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1), sig_a_sq = 10,
                     sig_b_sq = 10, error_sq = 1)
optimal_design
```

They can then score the design using either the A_crit or D_crit function.

```{r}
A_crit(optimal_design)
D_crit(optimal_design)
```

